\documentclass[letterpaper, 10pt, notitlepage]{report}
\usepackage[plainpages=false, colorlinks, urlcolor=black, citecolor=black,
	linkcolor=black]{hyperref}
\usepackage{graphicx}

%A LaTeX package which provides macros for the graphical
%representation of the keys on a computer keyboard.
\usepackage{keystroke}

%http://ctan.dcc.uchile.cl/help/Catalogue/entries/floatrow.html
%\usepackage{floatrow}

\usepackage{longtable}

\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{fixltx2e}

\usepackage[table,xcdraw]{xcolor}
\usepackage{dcolumn}


\usepackage[utf8]{inputenc}
\usepackage{txfonts}
\usepackage{listings}
\renewcommand{\lstlistingname}{Pseudocódigo}
\usepackage{multirow}
%\usepackage{float}
\usepackage{tikz}
\usepackage[polutonikogreek,english,spanish]{babel}
\usepackage{epsfig}
\usepackage{./sty/utfsm_tesis}

%Include Table of Contents as the first entry in TOC
\usepackage{./sty/xtocinc}
\usepackage[nottoc]{tocbibind}

%\usepackage[dvips]{epsfig}
%\usepackage{psfig}

\usepackage{caption}
\captionsetup[table]{name=Tabla}
\usepackage{subcaption}
\usepackage{verbatim}

%% Margenes segun Normas
%% Ancho Legal 21,59cm /  8,5in
%% Alto  Legal 33,02cm / 13,0in
%\paperheight    27.81cm % alto letter
%\paperwidth     21.59cm % ancho
%\hoffset        -1.0in % Seteo a 0 el margen izquierdo
%\voffset        -1.0in % Seteo a 0 el margen superior
%\oddsidemargin  3.80cm % Margen izquierdo (pag. impar)
%% \evensidemargin 2.55cm % Margen izquierdo (pag. par) Acrobat Winkk
%\evensidemargin 2.59cm % Margen izquierdo (pag. par)
%\topmargin      1.00cm % Margen superior
%\headheight     5.00mm % Ancho encabezado
%\headsep        8.00mm % Separacion encabezado-cuerpo
%\textheight     23.5cm % Alto cuerpo
%\textwidth      15.2cm % Ancho cuerpo
%\footskip       1.30cm % Separacion piepag-cuerpo
\parindent      0em
\parskip        2ex

% Fuzz -------------------------------------------------------------------
\hfuzz2pt

% \oddsidemargin	 0cm	% Ancho Legal 21,59cm
% \evensidemargin 0.5cm	% Alto	Legal 35,56cm
% \textwidth	 16.5cm
% \topmargin	  -1.5cm
% \textheight	 22cm

\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

\newcommand{\setlinespacing}[1]%
	   {\setlength{\baselineskip}{#1 \defbaselineskip}}
\newcommand{\doublespacing}{\setlength{\baselineskip}%
			   {1.3 \defbaselineskip}}
\newcommand{\singlespacing}{\setlength{\baselineskip}{\defbaselineskip}}


\lstloadlanguages{C++, sh, IDL, make}
\lstset{basicstyle=\small\sffamily, commentstyle=\slshape,
        numbers=left, numberstyle=\tiny, numbersep=10pt,
        extendedchars, frame=lines,
        floatplacement=ht, captionpos=b,
        defaultdialect=[CORBA]IDL}



%descomentar para poner fecha distinta a fecha de compilacion
\copyrightyear{2018} \submitdate{Enero 2018}
\convocation{Enero}{2018}

\title{Algoritmo para el cálculo de fragmentos de proteínas en los organismos secuenciados}
\author{Felipe Nicolás Araya Barrera}

\begin{document}

\selectlanguage{spanish}

\profguia{Lioubov Dombrovskaia}
\profcorr{Diego Arroyuelo Billiardi y Alexander Zamyatnin}


% Archivo de Agradecimientos
\ack{include/acknowledgements}

% Incluir Resumen
\resumenesp{include/resumen}

% Incluir Abstract
\resumening{include/abstract}

% Incluir Abreviaciones
\abreviaciones{include/glossary}

\beforepreface
\afterpreface

%\numberwithin{equation}{chapter}

% Incluir introduccion
\input{include/introduccion}

%inicio desarrollo
%archivo definicion del problema
\input{include/definicion}
%archivo estado del arte
\input{include/estado_arte}
%archivo propuesta
\input{include/propuesta}
%archivo implementacion
\input{include/implementacion}
%fin desarrollo

%archivo conclusiones
\input{include/conclusiones}

\input{include/apendice}

\begin{thebibliography}{9}

\bibitem{bioinformatica} 
Lahoz-Beltrá, Rafael.
\textit{BIOINFORMÁTICA, simulación, vida artificial e inteligencia artificial}. 
Ediciones Díaz de Santos S.A., Madrid, España, 2004.

\bibitem{searching} 
Orcutt, Bruce; Barker, Winona.
\textit{Searching the protein sequence database}.
National Biomedical Research Foundation, Georgetown University Medical Center,
Washington, D.C., \textit{Bulletin of Mathematical Biology}, 46(4): 545-552, 1984.

\bibitem{array} 
Zamyatnin, Alexander.
\textit{The Features of an Array of Natural Oligopeptides}.
Bakh Institute of Biochemistry, Russian Academy of Sciences, Moscow, Russia;
Departamento de Informática, Universidad Técnica Federico Santa María, Valparaíso, Chile,
\textit{Neurochemical Journal}, 10(4): 249-257, 2016.

\bibitem{biomolecula} 
Wikipedia, Definición de biomolécula.
\\\texttt{https://es.wikipedia.org/wiki/Biomolécula}.

\bibitem{amino} 
Koolman, Jan; Röhm, Klaus-Heinrich.
\textit{Bioquímica: Texto y Atlas}.
Editorial Médica Panamericana, Madrid, España;
Georg Thieme Verlag, Stuttgart, Germany, 2004.

\bibitem{zamyatnin1} 
Zamyatnin, Alexander.
\textit{Fragmentomics of Natural Peptide Structures}. 
Bach Institute of Biochemistry, Russian Academy of Sciences; Universidad Técnica Federico Santa María, Departamento de Informática, Valparaíso.
\textit{Biochemistry (Moscow)}, 74(13): 1575-1585, 2009.

\bibitem{zamyatnin2} 
Zamyatnin, Alexander.
\textit{Fragmentomics of Oligopeptides and Proteins}. 
4th International Peptide Symposium in conjunction with 7th Australian Peptide Symposium and 2nd Asia-Pacific International Peptide Symposium. Cairns, Australia, Conference Program and Abstracts, p. 130, 2007.

\bibitem{fuerzabruta} 
Búsqueda secuencial de texto, Algoritmo Fuerza Bruta.
\\\texttt{https://sites.google.com/site/busquedasecuencialdetexto/home}.

\bibitem{stringmatching} 
Navarro, Gonzalo; Raffinot, Mathieu.
\textit{Flexible Pattern Matching Strings: Practical On-Line Search Algorithms for Texts and Biological Sequences}.
Universidad de Chile; Centre Nationale de Recherche Scientifique, Marne-La-Vallee, Francia.
\textit{Cambridge University Press}, Nueva York, Estados Unidos, 2002.

\bibitem{knuthmorrispratt} 
Knuth, Donald; Morris, James y Pratt, Vaughan.
\textit{Fast pattern matching in strings}.
\textit{SIAM Journal on Computing}, 6(2): 323-350, 1977.

\bibitem{boyermoore} 
Boyer, Robert; Moore, Strother.
\textit{A Fast String Searching Algorithm}.
\textit{Communications of the ACM}, Nueva York, Estados Unidos, 20(10): 762-772, 1977.

\bibitem{pekka} 
Kilpelanen, Pekka.
\textit{Lecture 8: Applications of Suffix Trees}. 
University of Kuopio, Finland; Department of Computer Science, Spring 2005, pp. 1-20.


\bibitem{sufijos} 
\textit{Introducción a la biología computacional, árboles de sufijos}.
\\\texttt{ http://webdiis.unizar.es/asignaturas/TAP/material/4.2.sufijos.pdf}, pp. 1-16.

 
\bibitem{pavesi} 
Pavesi, Giulio; Mauri, Giancarlo; Pesole, Graziano.
\textit{An algorithm for finding signals of unknown length in DNA sequences}. 
Department of Computer Science, Systems and Communication, University of Milan-Bicocca, Via Bicocca degli Ancimboldi; Department of General Phisiology and Biochemistry, University of Milan, Via Celoria.
\textit{BIOINFORMATICS}, 17: 207-214, April 3rd, 2001.

\bibitem{suffixarray1} 
Manber, Udi; Myers, Eugene.
\textit{Suffix Arrays: A New Method for On-Line String Searches}. San Francisco, California, USA, 
\textit{Proceedings of the First Annual ACM-SIAM Symposium on Discrete Algorithms}, 319-327, 1990.

\bibitem{licenciado} 
Deymonnaz, Alejandro.
\textit{Arreglos de sufijos para alineamiento de secuencias de ADN con memoria acotada}. Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Computación, Argentina, 2012.

\bibitem{koaluru} 
Ko, Pang; Aluru, Srinivas.
\textit{Suffix Tree Applications in Computational Biology}. 
Iowa State University, \textit{Handbook of Computational Molecular Biology}, Chapter 6, 2001.

\bibitem{abeliuk} 
Abeliuk Kimmerman, Andrés.
\textit{Árboles de sufijos comprimidos para textos altamente repetitivos}. Universidad de Chile, Facultad de Ciencias Físicas y Matemáticas, Departamento de Ciencias de la Computación, Chile, 2012.

\bibitem{kurtz} 
Kurtz, Stefan.
\textit{Reducing the space requirement of suffix trees}.
\textit{Software: Practice and Experience}, 29(13): 1149-1171, 1999.

\bibitem{suffixarray2} 
Vladu, Adrian; Negru\c{s}eri, Cosmin.
\textit{Suffix arrays - a programming contest approach}. Noviembre 2005.

\bibitem{lcparray} 
Arreglo LCP, \textit{LCP array}.
\\\texttt{https://en.wikipedia.org/wiki/LCP\_array}.

\bibitem{kasai} 
Kasai, T.; Lee, G.; Arimura, H.; Arikawa, S.; Park, K.
\textit{Linear-Time Longest-Common-Prefix Computation in Suffix Arrays and Its Applications}.
\textit{Combinatorial Pattern Matching: 12th Annual Symposium}, 2089: 181-192, 2001. 

\bibitem{aminoacids}
The Twenty Amino Acids.
\\\texttt{https://www.cryst.bbk.ac.uk/education/AminoAcid/the\_twenty.html}.

\bibitem{manipulatingstrings}
Manipulating Strings.
\\\texttt{http://teacher.buet.ac.bd/faizulbari/cse201/Strings.pdf}.

\bibitem{kasaimethod}
Kasai’s Algorithm for Construction of LCP array from Suffix Array.
\\\texttt{http://www.geeksforgeeks.org/\%C2\%AD\%C2\%ADkasais-algorithm-for-}
\\\texttt{construction-of-lcp-array-from-suffix-array/}.

\bibitem{tutorial}
Soulié, Juan.
\textit{C++ Language Tutorial}. Junio 2007.

\bibitem{queues}
\textit{Chapter 11: Priority Queues and Heaps}.
\\\texttt{https://web.engr.oregonstate.edu/$\sim$sinisa/courses/OSU/CS261/}
\\\texttt{CS261\_Textbook/Chapter11.pdf}. 

\bibitem{better} 
Dementiev, Roman; Kärkkäinen, Juha; Mehnert, Jens; Sanders, Peter.
\textit{Better external memory suffix array construction}.
\textit{Journal of Experimental Algorithmics}, Nueva York, Estados Unidos, 12: 1-24, 2008.

\bibitem{esais} 
Bingmann, Timo; Fischer, Johannes; Osipov, Vitaly.
\textit{Inducing Suffix and LCP Arrays in External Memory}.
KIT, Institute of Theoretical Informatics, Karlsruhe, Alemania.
\textit{Journal of Experimental Algorithmics}, Nueva York, Estados Unidos, 21: 1-15, 2016.

\bibitem{sascan} 
Kärkkäinen, Juha; Kempa, Dominik.
\textit{Engineering a Lightweight External Memory Suffix Array Construction Algorithm}.
Department of Computer Science, University of Helsinki, Finlandia.
\textit{Mathematics in Computer Science}, 11(2): 137-149, 2017.

\bibitem{newindices} 
Gonnet, Gaston; Baeza-Yates, Ricardo; Snider, Tim.
\textit{New indices for text: PAT Trees and PAT arrays}.
\textit{Information Retrieval: Algorithms and Data Structures}, Prentice-Hall, Englewood Cliffs, 66-82, 1992.

\bibitem{bwt}
Burrows, Michael; Wheeler, David.
\textit{A block sorting lossless data compression algorithm}. 
Palo Alto, California, 1994.

\bibitem{bowtie} 
Langmead, Ben; Trapnell, Cole; Pop, Mihai; Salzberg, Steven.
\textit{Ultrafast and memory-efficient alignment of short DNA sequences to the human genome}. 
Center for Bioinformatics and Computational Biology, Institute for Advanced Computer Studies, University of Maryland, College Park, MD 20742, USA; Genome Biology, March 2009.

\bibitem{aplicacion}
Ohlebusch, Enno.
\textit{Bioinformatics Algorithms: Sequence Analysis, Genome Rearrangements, and Phylogenetic Reconstruction}. 
Oldenbusch Verlag, 2013.

\bibitem{lcpscan}
Kärkkäinen, Juha; Kempa, Dominik.
\textit{LCP array construction in external memory}.
Department of Computer Science, University of Helsinki, Finlandia.
\textit{Proceedings of the $13^{th}$ International Symposium on Experimental Algorithms (SEA 2014)}.
8504: 412-423, Springer, 2014.

\bibitem{emsparse}
Kärkkäinen, Juha; Kempa, Dominik.
\textit{Faster External Memory LCP Array Construction}.
Department of Computer Science, University of Helsinki, Finlandia.
\textit{Proceedings of the $24^{th}$ Annual European Symposium on Algorithms (ESA 2016)}.
61: 61:1-61:16, Schloss Dagstuhl–Leibniz-Zentrum fuer Informatik, 2016.

\bibitem{emsparse2}
Kärkkäinen, Juha; Kempa, Dominik.
\textit{Engineering External Memory LCP Array Construction: Parallel, In-Place and Large Alphabet}.
Department of Computer Science, University of Helsinki, Finlandia.
\textit{Proceedings of the $16^{th}$ International Symposium on Experimental Algorithms (SEA 2017).}.
17: 17:1-17:14, Schloss Dagstuhl–Leibniz-Zentrum fuer Informatik, 2017.

\bibitem{codeblocks}
Goetz, Lawrence; Langsam, Yedidyah; Raphan, Theodore.
\textit{Code::Blocks Student Manual}.
Department of Computer and Information Science, Brooklyn College of CUNY.
Version 16.01, 2016.

\bibitem{navarro2016}
Navarro, Gonzalo.
\textit{Compact Data Structures: A Practical Approach}.
Departamento de Ciencias de la Computación, Universidad de Chile.
\textit{Cambridge University Press}, Nueva York, Estados Unidos, 2016.

\bibitem{quimicaaminoacidos}
Naturaleza Química de los Aminoácidos.
\\\texttt{https://themedicalbiochemistrypage.org/es/amino-acids-sp.php}.

\bibitem{zamyatnin3}
Zamyatnin, Alexander. 
\textit{Hemoglobin as a potential source of natural regulatory oligopeptides}. 
Bach Institute of Biochemistry, Russian Academy of Sciences; Universidad Técnica Federico Santa María, Departamento de Informática, Valparaíso. 
\textit{Biochemistry (Moscow)}, 74(2): 201-208, 2009.

\bibitem{vbyte}
Williams, Hugh; Zobel, Justin.
\textit{Compressing integers for fast file access}.
Department of Computer Science, RMIT University, Australia.
\textit{The Computer Journal}, 42(3): 193-201, 1999.

\bibitem{erop2}
Zamyatnin, Alexander. 
\textit{EROP-Moscow: specialized data bank for endogenous regulatory oligopeptides}. 
Anokhin Institute of Normal Physiology, USSR Academy of Medical Sciences. 
\textit{Protein Sequence and Data Analysis}, 4(1): 49-52, 1991.

\bibitem{erop3}
Zamyatnin, Alexander; Borchikov, Alexander; Vladimirov, Mikhail; Voronina, Olga. 
\textit{The EROP-Moscow oligopeptide database}. 
A.N.Bakh Institute of Biochemistry, Russian Academy of Sciences. 
\textit{Nucleic Acids Research}, 34: D261-D266, 2006.


\bibitem{swissprot}
UniProt-SwissProt database.
\\\texttt{http://www.uniprot.org/uniprot/?query=reviewed:yes}.

\bibitem{trembl}
UniProt-TrEMBL database.
\\\texttt{http://www.uniprot.org/uniprot/?query=reviewed:no}.

\bibitem{eropmoscow}
EROP-Moscow: Endogenous Regulatory OligoPeptide knowledgebase.
\\\texttt{http://erop.inbi.ras.ru/}.

\bibitem{fasta}
Wikipedia, Definición de archivo FASTA.
\\\texttt{https://es.wikipedia.org/wiki/Formato\_FASTA}.

\bibitem{yutamori}
Yuta Mori. libdivsufsort, a C library for suffix array construction.
\\\texttt{http://code.google.com/p/libdivsufsort/}.

\end{thebibliography}

%\bibliographystyle{plain}
\bibliography{bib/papers}


%\backmatter
%\cleardoublepage

\singlespacing
\cleardoublepage





\end{document}
